The ChEMBL database as Linked Open Data

### 2012-07-09 article.tex
Tuned more results

The ChEMBL database
recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
explicit types.

@@

For drug discovery, the drugs themselves are the main topic of study.
ChEMBL contains many different drug types, mostly small molecules,
but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
even cells. All compounds are represented as classes, following
the design of the CHEMINF ontology. These entities do not have a common
superclass in common, but can easily be identified as having the 
the role of being a drug. This is triplified using the OBO and ChEBI
ontologies:

@@

The entity typing is expressed
by subclassing other classes. For example, proteins subclass the
Protein concept from the PRotein Ontology (PR\_000000001), small
molecules subclass the Chemical Entity concept from the CHEMINF
ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
subclass their respective matches in the CHEBI ontology (CHEBI\_50699
and CHEBI\_7754). For each drug compound the name and synonyms are
provided as labels. When InChI and InChIKeys are available for a drug, then these are
provided via the CHEMINF formalism (here abbreviated):

@@

For small molecules, molecular properties are often available from ChEMBL, and as of ChEMBL-RDF 13
these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
CHEMINF ontology approach. Here is, for example, the ALogP value:

@@

For documents little information is replicated from the database, taking advantage
of PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs). URIs for the
latter can be resolved online, providing curated information on the identified
documents. For each paper, basic properties are provided using the BIBO
ontology:

\begin{small}
\begin{verbatim}
journal:j6c706049c2e08871b7c46a6528065736
 a bibo:Journal ;
 dc:title "J. Med. Chem." .

res:r1
 a bibo:Article ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .
\end{verbatim}
\end{small}

>>>>

The ChEMBL database
recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
explicit types, and effectively uses a closed-world approach. A target
specification can then look like:

@@

For drug discovery, the drugs themselves are the main topic of study.
ChEMBL contains many different drug types, mostly small molecules,
but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
even cells. ChEMBL-RDF follows the approach used by the CHEMINF ontology,
and encodes drugs as classes, rather than instances, and they subclass
classes defined by several ontologies for these drug types. For example, proteins subclass the
Protein concept from the PRotein Ontology (PR\_000000001), small
molecules subclass the Chemical Entity concept from the CHEMINF
ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
subclass their respective matches in the CHEBI ontology (CHEBI\_50699
and CHEBI\_7754).

It should also be noted that the drugs are not subclassed from
a general drug class, or any other common superclass. However, to annotate
the drugs to be drugs, they are identified as having the 
the role of being a drug. This is triplified using the OBO and ChEBI
ontologies, where CHEBI\_23888 is the drug role:

@@

The name and synonyms for each drug are
provided as labels. When InChI and InChIKeys are available for a drug, then these are
provided via the CHEMINF formalism (here abbreviated):

@@

Similarly, properties of small molecules are available from ChEMBL, and as of ChEMBL-RDF 13
these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
CHEMINF ontology approach. Here is, for example, the ALogP value:

@@

The literature from which all this data was extracted, is replicated from the database,
but only a subset of information about that literature is provided. In all cases,
it uses the PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs) to link
to authoritative databases to provide full detail, as outlined later in this paper.
The provided properties are using the BIBO ontology:

\begin{small}
\begin{verbatim}
res:r1
 a bibo:Article ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .

journal:j6c706049c2e08871b7c46a6528065736
 a bibo:Journal ;
 dc:title "J. Med. Chem." .
\end{verbatim}
\end{small}
